---
title: "02_DataUnderstanding"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```


```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(ProjectTemplate)
load.project()

```

# Data Understanding

The datasets was created during a run using 1024 GPU nodes. There are 3 levels visualisation output; 4, 8, 12.

The data provided shows performance timing of the render application, performance of the GPU card, and details of which part of the image was being rendered in each task.


## application-checkpoints.csv (111.2MB)

This file contains application checkpoint events throughout the execution of the render job. Examples of the events are given for the `eventName` field.

### Schema

1. **timestamp:**    time of data creation. Ex: `2018-11-08T07:41:55.921Z`
2. **hostname:**   unique hostname created by the Azure batch system. Ex: `0d56a730076643d585f77e00d2d8521a00000N`
3. **eventName** Name of the event occuring within the rendering application.
    + **TotalRender** is the entire task
    + **Render** is when the image tile is is being rendered
    + **Saving Config** is simply a measure of configuration overhead
    + **Tiling** is where post processing of the rendered tile is taking place
    + **Uploading** is where the output from post processing is uploaded to Azure Blob Storage

4. **eventType:** 
    + **START**
    + **STOP**

5. **jobId:**    ID of the Azure batch job. Ex: `1024-lvl12-7e026be3-5fd0-48ee-b7d1-abd61f747705`

6. **taskId:**   ID of the Azure batch task. Ex: `b47f0263-ba1c-48a7-8d29-4bf021b72043`



### Summary 

```{r,  echo=FALSE}
summary(application.checkpoints)
```


## gpu.csv (208.7MB)

This file contains metrics that were output regarding the status of the GPU on the virtual machine.

## Schema

1. **timestamp:**  time of data creation. Ex: `2018-11-08T08:27:10.314Z`

2. **hostname:** unique hostname created by the Azure. Ex: `8b6a0eebc87b4cb2b0539e81075191b900001C`

3. **gpuSerial:**  The serial number of the physical GPU card.Ex: "0323217055910"

4. **gpuUUID:** The unique system id assigned by the Azure system to the GPU unit. Ex: `GPU-1d1602dc-f615-a7c7-ab53-fb4a7a479534`

5. **powerDrawWatt:**    Power draw of the GPU in watts. Ex: `131.55`

6. **gpuTempC:**   Temperature of the GPU in Celsius Ex: `48`

7. **gpuUtilPerc:**    Percent utilisation of the GPU Core(s). Ex: `92`

8. **gpuMemUtilPerc:**   Percent utilisation of the GPU memory. Ex: `52`


### Summary 

```{r,  echo=FALSE}
summary(gpu)
```


---

## task-x-y.csv (6.2MB)

This file contains the x,y co-ordinates of which part the image was being rendered for each task.

### Schema

1. **jobId:** Id of the Azure batch job. Ex: `1024-lvl12-7e026be3-5fd0-48ee-b7d1-abd61f747705`

2. **taskId:**   ID of the Azure batch task. Ex: `b47f0263-ba1c-48a7-8d29-4bf021b72043`

3. **x:**  X co-ordinate of the image tile being rendered. Ex: `116`

4. **y:**    Y co-ordinate of the image tile being rendered. Ex: `178`

5. **level:** Zoom level
    + **4**
    + **8**
    + **12**


### Summary 

```{r,  echo=FALSE}
summary(gpu)
```

---




